Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-10-14 Simulation.tex
edits to thesis from clarice and modify background to reflect bistatic

\section{Introduction}
The following simulations performed will define the search space on the effect of rotor blade modulation. Since the software developed allows for full parameterization of the objects within the scene the simulations will focus on a few that effect the different object's positions in 3D space and the resulting received signal. The parameters that will be changed are the position of the receiver, pitch of the rotor blade, and transmitter position. The effect of each parameter will be analyzed to determine what information can be derived for localization.

@@

\section{Analysis Techniques}
The data from the .CSV files generated by the raytracing program is loaded into MATLAB scripts to process the data. The technique that is used is the spectrogram function that computes the short time Fourier transform. This allows for computations to be done in the time frequency domain and provide an accurate picture of how the signal is changing with time. The following figure \label{fig:test_spec} shows an example of the computed spectrograms.

@@

After the computation of the short time Fourier transform the doppler envelope is calculated by performing a threshold on the power, in both the positive and negative doppler, at each time slice. The envelope value for that time slice is found by approaching zero frequency and stopping as the power reaches a predetermined level from both the negative and positive values. The power level to set the threshold is calculated differently based on weather the elevation or the azimuth parameter is being swept. For the azimuth angle sweep, which will be shown later in section \ref{taa}, the maximum power is found in each time slice then those values are averaged. Then for each successive angle the same average is taken and averaged with all the others

@@

This technique is used because the elevation angle is remaining constant, so the distance with the helicopter platform should be constant as well. When varying the elevation angle in section \ref{sec:tea} the same method cannot be used because the distance is changing so only the maximum power is found in each time slice then those values are averaged without any successive averaging.

@@

The Doppler envelopes are displayed in figure \ref{fig:test_spec_w_doppler_profile}, the lower in blue and the upper in red. The minimum is calculated from the lower envelope, and the maximum from the upper envelope. Those values are then used to characterize that doppler profile and analyze signal trends as different variables change.

@@

The rotor blade length from figure \ref{fig:rx_position_image} is 7.5m so the receiver will be varied from 0m to 9m in 1m increments. The resulting signal is analyzed by calculating its doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of the position of the receiver.

@@

The figure \ref{fig:rx_position} shows maximum doppler vs the position of the receiver. From this graph the maximum doppler frequency changes significantly as the receiver is moved from directly under neath the rotor to past the tip of the rotor at 9m.The position of the receiver also effects the resulting signal envelope shapes shown in figures \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m}.

@@

Comparing figure \ref{fig:rx_position_0m} with \ref{fig:rx_position_7m} the max doppler reflects the information in figure \ref{fig:rx_position} and the overall envelope is afected. The envelope in figure \ref{fig:rx_position_0m} is much noiser than the envelope of figure \ref{fig:rx_position_7m} and the envelope of this second figure has a more distinct shape with less variation between its two peaks. The position of the receiver will also determine whether the azimuth angle can be derived from the resulting signal shown later in section \ref{taa}. Therefore the receiver will be offset from the center of rotation by 7m for the subsequent simulations.

@@

The resulting signal is analyzed by calculating its doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of pitching the rotor blade.

@@

From figure \ref{fig:pitch_15_135deg} the effect of pitch on the maximum doppler frequency is a linear increase between a pitch from 0 \textdegree \space to 15\textdegree. Due to the linear increase the pitch of the rotor blade can be accounted for when making measurements on subsequent simulation analysis. One issue does arise in when the transmitter is located underneath the helicopter which is shown in figure \ref{fig:pitch_tx0}.

@@

As the pitch increases linearly at first then climbs at a much faster rate. This offset when the transmitter is directly underneath is caused by the pitch of the blade reflecting rays into the receiver which would be reflected back at the ground if the pitch was 0\textdegree. If the assumption is that the helicopter is hovering when these measurements take place the pitch of the rotor will be much less than the maximum pitch available. This assumption will adjust the maximum pitch to 15 \textdegree \space which will allow for the linear assumption to be fulfilled even when the transmitter is directly underneath the helicopter.

@@

In figure \ref{fig:pitch_tx_elevation_angle} the max doppler vs elevation angles are initially shifted up in frequency from a pitch of 0 and has changed in slope. This is because the rays hitting the blades from below are reflecting into the receiver at an angle that is closer to the direction of blade travel. The interesting thing about this plot is that the pitch envelopes actually become closer to the 0\textdegree \space envelope as the elevation angle becomes smaller. Figure \ref{fig:pitch_tx_elevation_angle_difference} shows the difference plotted over elevation angle.

@@

From figure \ref{fig:pitch_tx_elevation_angle_difference} we can see that the difference in frequency changes linearly as the elevation angle decreases, the slope of which is based on the pitch angle. This slope would have to be characterized on the physical craft and can be accounted for due to its predictable nature.

@@

Where $\theta_{Az}$ is the azimuth angle. The azimuth angle of the transmitter in the simulation will effect the amount of doppler added to the signal. This is due to the direction of the rotor blade rotation as it approaches both the receiver and the transmitter and how the rays reflect off the blade as it moves. From the model we see the amount of doppler form a sinusoidal pattern as the azimuth angle goes from $0$ to $2\pi$. This is mirrored in the simulation shown in figure \ref{fig:tx_azimuth_angle_200}.

@@

The graph in figure \ref{fig:tx_azimuth_angle_200} shows both the max and the minimum measured doppler as the azimuth angle completes a full rotation. The doppler frequency follows the azimuth angle in a sinusoidal fashion. The sinusoidal pattern is held as the elevation angle changes shown in figures \ref{fig:tx_azimuth_angle_50} and \ref{fig:tx_azimuth_angle_400}.

@@

%description
The sinusoidal pattern of min and max doppler frequencies are only present when the receiver is located away from the axis of rotation. This is shown in figure \ref{fig:tx_azimuth_rx0} which is devoid of any sinusoidal pattern and remains constant as the transmitter azimuth angle is changed.

@@

The effect of pitching the blade is that it modifies the results of the azimuth sweep when compared to figure \ref{fig:tx_azimuth_angle_200}. The resulting min and max envelope frequencies vs azimuth angle for pitch are shown in \ref{fig:pitch_azimuth_rev}. This occurs because the pitch of the blade angles RF reflections more in the direction of travel but not at all points in the azimuth. When the transmitter is located between 90\textdegree and 270\textdegree pitch causes an increase in doppler but elsewhere decreases the amount of doppler shift. The decrease is caused by the blade surface top reflecting receding doppler away from the receiver at those angles.

@@

The effect of azimuth angle on the maximum measured doppler will be characterized in section \ref{sec:transmitter_azimuth_angle_estimation}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------------
\section{Transmitter Elevation Angle} \label{sec:tea}
%display elevation angle with respect to max doppler readings and describe patterns
The transmitter elevation angle is the angle between the helicopter rotor and the position of the transmitter shown in figure \ref{fig:azimuth_rel}.

@@

Where $\theta_{El}$ is the elevation angle.

The smaller elevation angles mean that the transmitter is farther away from the helicopter. The elevation angle in relation to the distance of the transmitter is shown by figure \ref{fig:tx_range_elevation_rel}.

@@

The elevation angle of the transmitter in the simulation will effect the amount of doppler added to the signal. This is due to the rays having a vector component in the direction of the travel of the rotor. When the transmitter is directly underneath the rotor and receiver there is no vector component in the direction of rotor travel so the reflected signal is not doppler shifted. From the model we see the amount of doppler increase as the elevation angle decreases. This is mirrored in the simulation shown in figure \ref{fig:tx_elevation_135deg}.

@@

The graph in figure \ref{fig:tx_elevation_135deg} shows the maximum measured doppler as the elevation angle increases.

Looking back at the geometric models for transmitter azimuth angles of 0\textdegree \space, 90\textdegree \space, 180\textdegree \space, and 270\textdegree \space and their simulated counterparts shown in figure \ref{fig:verify}.

@@

The effect of elevation angle on the maximum measured doppler will be characterized in section \ref{sec:transmitter_elevation_angle_estimation}.

>>>>

\section{Introduction}
The following simulations performed will define the search space on the effect of rotor blade modulation. Since the software developed allows for full parameterization of the objects within the scene, the simulations will focus on a few that effect the different object's positions in 3D space and the resulting received signal. The parameters that will be changed are the position of the receiver, pitch of the rotor blade, and transmitter position. The effect of each parameter will be analyzed to determine what information can be derived for localization.

@@

\section{Analysis Techniques}
The data from the CSV files generated by the raytracing program are loaded into MATLAB scripts to process the data. The processing technique used is the spectrogram function that computes the short time Fourier transform. This allows for computations to be done in the time frequency domain and provides an accurate picture of how the signal is changing with time. Figure \label{fig:test_spec} shows an example of the computed spectrograms.

@@

After the computation of the short time Fourier transform; the Doppler envelope is calculated by performing a threshold on the power, in both the positive and negative Doppler, at each time slice. The envelope value for that time slice is found by approaching zero frequency and stopping as the power reaches a predetermined level from both the negative and positive values. The power level to set the threshold is calculated differently based on weather the elevation or the azimuth parameter is being swept. For the azimuth angle sweep, which will be shown later in section \ref{taa}, the maximum power is found in each time slice then those values are averaged. Then for each successive angle the same average is taken and averaged with all the others.

@@

This technique is used because the elevation angle is remaining constant, so the distance with the helicopter platform should be constant as well. When varying the elevation angle in section \ref{sec:tea}, the same method cannot be used because the distance is changing so only the maximum power is found in each time slice. Then all those values are averaged without any successive averaging.

@@

The Doppler envelopes are displayed in Figure \ref{fig:test_spec_w_doppler_profile}, the lower in blue and the upper in red. The minimum is calculated from the lower envelope, and the maximum from the upper envelope. Those values are then used to characterize that Doppler profile and analyze signal trends as different variables change.

@@

The rotor blade length from Figure \ref{fig:rx_position_image} is 7.5m so the receiver will be varied from 0m to 9m in 1m increments. The resulting signal is analyzed by calculating its Doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of the position of the receiver.

@@

Figure \ref{fig:rx_position} shows maximum Doppler vs the position of the receiver. From this graph the maximum Doppler frequency changes significantly as the receiver is moved from directly under neath the rotor to past the tip of the rotor at 9m.The position of the receiver also effects the resulting signal envelope shapes shown in Figures \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m}.

@@

Comparing Figure \ref{fig:rx_position_0m} with \ref{fig:rx_position_7m} the max Doppler reflects the information in Figure \ref{fig:rx_position} and the overall envelope is afected. The envelope in Figure \ref{fig:rx_position_0m} is much noiser than the envelope of Figure \ref{fig:rx_position_7m} and the envelope of this second figure has a more distinct shape with less variation between its two peaks. The position of the receiver will also determine whether the azimuth angle can be derived from the resulting signal shown later in section \ref{taa}. Therefore the receiver will be offset from the center of rotation by 7m for the subsequent simulations.

@@

The resulting signal is analyzed by calculating its Doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of pitching the rotor blade.

@@

From Figure \ref{fig:pitch_15_135deg} the effect of pitch on the maximum Doppler frequency is a linear increase between a pitch from 0 \textdegree \space to 15\textdegree. Due to the linear increase the pitch of the rotor blade can be accounted for when making measurements on subsequent simulation analysis. One issue does arise in when the transmitter is located underneath the helicopter which is shown in Figure \ref{fig:pitch_tx0}.

@@

The pitch increases linearly at first then climbs at a much faster rate. This offset when the transmitter is directly underneath is caused by the pitch of the blade reflecting rays into the receiver which would be reflected back at the ground if the pitch was 0\textdegree. If the assumption is that the helicopter is hovering when these measurements take place the pitch of the rotor will be much less than the maximum pitch available. This assumption will adjust the maximum pitch to 15\textdegree \space which will allow for the linear assumption to be fulfilled even when the transmitter is directly underneath the helicopter.

@@

In Figure \ref{fig:pitch_tx_elevation_angle} the max Doppler vs elevation angles are initially shifted up in frequency from a pitch of 0\textdegree \space and has changed in slope. This is because the rays hitting the blades from below are reflecting into the receiver at an angle that is closer to the direction of blade travel. The interesting thing about this plot is that the pitch envelopes actually become closer to the 0\textdegree \space envelope as the elevation angle becomes smaller. Figure \ref{fig:pitch_tx_elevation_angle_difference} shows the difference plotted over elevation angle.

@@

From Figure \ref{fig:pitch_tx_elevation_angle_difference} we can see that the difference in frequency changes linearly as the elevation angle decreases, the slope of which is based on the pitch angle. This slope would have to be characterized on the physical craft and can be accounted for due to its predictable nature.

@@

Where $\theta_{Az}$ is the azimuth angle. The azimuth angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the direction of the rotor blade rotation as it approaches both the receiver and the transmitter and how the rays reflect off the blade as it moves. From the model we see the amount of Doppler form a sinusoidal pattern as the azimuth angle goes from $0$ to $2\pi$. This is mirrored in the simulation shown in Figure \ref{fig:tx_azimuth_angle_200}.

@@

The graph in Figure \ref{fig:tx_azimuth_angle_200} shows both the max and the minimum measured Doppler as the azimuth angle completes a full rotation. The Doppler frequency follows the azimuth angle in a sinusoidal fashion. The sinusoidal pattern is held as the elevation angle changes shown in Figures \ref{fig:tx_azimuth_angle_50} and \ref{fig:tx_azimuth_angle_400}.

@@

%description
The sinusoidal pattern of min and max Doppler frequencies are only present when the receiver is located away from the axis of rotation. This is shown in Figure \ref{fig:tx_azimuth_rx0} which is devoid of any sinusoidal pattern and remains constant as the transmitter azimuth angle is changed.

@@

Pitching the blade modifies the results of the azimuth sweep when compared to Figure \ref{fig:tx_azimuth_angle_200}. The resulting min and max envelope frequencies vs azimuth angle for pitch are shown in \ref{fig:pitch_azimuth_rev}. This occurs because the pitch of the blade angles RF reflections differently according to transmitter azimuth. When the transmitter is located between 90\textdegree \space and 270\textdegree \space pitch causes an increase in Doppler but elsewhere decreases the amount of Doppler shift. The decrease is caused by the blade surface top reflecting receding Doppler away from the receiver at those angles.

@@

The effect of azimuth angle on the maximum measured Doppler is characterized in section \ref{sec:transmitter_azimuth_angle_estimation}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------------
\section{Transmitter Elevation Angle} \label{sec:tea}
%display elevation angle with respect to max doppler readings and describe patterns
The transmitter elevation angle is the angle between the helicopter rotor and the position of the transmitter shown in Figure \ref{fig:azimuth_rel} where $\theta_{El}$ is the elevation angle.

@@

The smaller elevation angles mean that the transmitter is farther away from the helicopter. The elevation angle in relation to the distance of the transmitter is shown by Figure \ref{fig:tx_range_elevation_rel}.

@@

The elevation angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the rays having a vector component in the direction of the travel of the rotor. When the transmitter is directly underneath the rotor and receiver there is no vector component in the direction of rotor travel so the reflected signal is not Doppler shifted. From the model we see the amount of Doppler increase as the elevation angle decreases. This is mirrored in the simulation shown in Figure \ref{fig:tx_elevation_135deg} where the maximum measured Doppler as the elevation angle increases.

@@

Looking back at the geometric models for transmitter azimuth angles of 0\textdegree \space \ref{fig:3D_model_0az_doppler}, 90\textdegree \space \ref{fig:3D_model_90az_doppler}, 180\textdegree \space \ref{fig:3D_model_180az_doppler}, and 270\textdegree \space \ref{fig:3D_model_270az_doppler} and their simulated counterparts shown in Figure \ref{fig:verify}. We see that the ray traced results match their projections on the maximum and minimum envelope Doppler as transmitter elevation angle changes. The Doppler values do not match exactly but the trends are very similar.

@@

The effect of elevation angle on the maximum measured Doppler will be characterized in section \ref{sec:transmitter_elevation_angle_estimation}.
